---
title: "Homework 3"
output: github_document
Author: Adam R
editor_options: 
  chunk_output_type: console
---
```{r setup, include= FALSE}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
```

## Problem 1

```{r}
data("instacart")
```

this dataset contains `r nrow(instacart)` and `r ncol(instacart)` columns. 

Observations are the level of items in order by user. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. [ADD MORE BEFORE SUBMITTING].

how many aisles, and which are they from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles, and the most ordered are fresh vegetables, fresh fruit, and packaged vegetables and fruits.

let's make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

making a table

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

apples v ice cream table 
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

## Problem 2




## Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```


